---
title: "README"
author: "Abhishek salape"
date: "12/06/2020"
output: html_document
---
##Assignment Instructions  
1.Code for reading in the dataset and/or processing the data  
2.Histogram of the total number of steps taken each day  
3.Mean and median number of steps taken each day  
4.Time series plot of the average number of steps taken  
5.The 5-minute interval that, on average, contains the maximum number of steps  
6.Code to describe and show a strategy for imputing missing data  
7.Histogram of the total number of steps taken each day after missing values are imputed  
8.Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends  
9.All of the R code needed to reproduce the results (numbers, plots, etc.) in the report  

**STEP 1**  
##code for Reading & Exploring The data 
```{r,echo=TRUE,results="hide"}

setwd("C:/Users/Hp/Documents")
activity <- read.csv("./activity.csv")
head(activity)
str(activity)
dim(activity)
sum(is.na(activity$steps))

  ```
  
**STEP 2**  
##Histogram of the total number of steps taken each day
```{r,echo=TRUE,results='hide'}
  
library(ggplot2)
setwd("C:/Users/Hp/Documents")
activity <- read.csv("./activity.csv")
Q2<-data.frame(tapply(activity$steps,activity$date,sum,na.rm=TRUE))
Q2$date<-rownames(Q2)
rownames(Q2)<-NULL
names(Q2)[[1]]<-"Total Steps"
#Total Steps by date bar chart
png("plot11.png")
ggplot(Q2,aes(y=Q2$`Total Steps`,x=Q2$date))+geom_bar(stat="identity") + ylab("Total Steps")+xlab("Date")+ggtitle("Total Steps by date")
dev.off()

```

**STEP 3**  
##Mean and median number of steps taken each day

```{r,echo=TRUE,results='hide'}
setwd("C:/Users/Hp/Documents")
activity <- read.csv("./activity.csv")
totalStepsByDay<-aggregate(steps~date, activity, sum)
mean(totalStepsByDay$steps)
median(totalStepsByDay$steps)

```
**STEP 4**
##Time series plot of the average number of steps taken  
```{r,echo=TRUE,results='hide'}
averageStepsbyInterval<-aggregate(steps~interval, activity, mean)
png("plot12.png")
with(averageStepsbyInterval, plot(interval, steps, type = "l"))  

```
**STEP 5**  
##Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?  

```{r,echo=TRUE,results='hide'}

averageStepsbyInterval[which.max(averageStepsbyInterval[,2]),1]

```


**STEP 6**
Code to describe and show a strategy for imputing missing data
There are multiple strategies to deal with multiple value imputations.
The common strategies include:
1. Constant value imputations
2. Regression model value imputations
3. Mean/mode value substitutions
For the purpose of simplicity, in this question, I will use the mean/mode value substitution strategy to impute missing values. That is, using the mean values to substitute out the missing values in the original data set
Before doing any sort of imputation, it is helpful to understand what are the distributions of missing values by date and interval
```{r, echo = TRUE,results="hide"}
Q6<-activity
Q6$Missing<-is.na(Q6$steps)
Q6<-aggregate(data=Q6,Missing~date+interval,FUN="sum")
Q6.1<-data.frame(tapply(Q6$Missing,Q6$date,sum))
Q6.1$date<-rownames(Q6.1)
rownames(Q6.1)<-NULL
names(Q6.1)<-c("Missing","date")
Q6.1$date<-as.Date(Q6.1$date,format="%Y-%m-%d")
Q6.2<-data.frame(tapply(Q6$Missing,Q6$interval,sum))
Q6.2$date<-rownames(Q6.2)
rownames(Q6.2)<-NULL
names(Q6.2)<-c("Missing","Interval")
par(mfrow=c(1,2))
png("plot15.png")
plot(y=Q6.1$Missing,x=Q6.1$date,main="Missing Value Distribution by Date")
png("plot16.png")
plot(y=Q6.2$Missing,x=Q6.2$Interval,main="Missing Value Distribution by Interval")
table(activity$date)
dev.off()
```

By this point, from the plot, that the missing values have a very disctinct pattern. For every interval, there are consistantly 8 missing values. For the date, there are consistantly 288 missing values. And in total, there are 8 dates that have missing value. We don't exactly know the cause for these missing values but there's a pattern. For that matter, we can see that the mean value imputation is appropriate.

We can see that every date has 288 data points. It means that the 8 dates have no data points at all what so ever. We can refine the analysis by looking at these missing values depending on their Weekday and interval parameters to matach with the average 

**STEP7**

Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.

```{r,echo=TRUE,results="hide"}
activity$date <- as.Date(strptime(activity$date, format="%Y-%m-%d"))
activity$datetype <- sapply(activity$date, function(x) {
        if (weekdays(x) == "Sábado" | weekdays(x) =="Domingo") 
                {y <- "Weekend"} else 
                {y <- "Weekday"}
                y
        })
```

Make a panel plot containing a time series plot (i.e. type = “l”) of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). See the README file in the GitHub repository to see an example of what this plot should look like using simulated data.

```{r,echo=TRUE,results="hide"}
activity_by_date <- aggregate(steps~interval + datetype, activity, mean, na.rm = TRUE)
png("plot17.png")
 ggplot(activity_by_date, aes(x = interval , y = steps, color = datetype)) +
       geom_line() +
       labs(title = "Average daily steps by type of date", x = "Interval", y = "Average number of steps") +
       facet_wrap(~datetype, ncol = 1, nrow=2)
 ```
























